SIR Model With Testing Flows and Isolation

### 2021-01-17 SIR_manuscript.tex
Faddy updated

Here, we develop a compartmental model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, reporting time and isolation on epidemic dynamic. We investigate the effect of underlying parameters of the model on the basic reproduction number, $\Rnum$. While targeted testing strategies are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is viral spread is expected to be stronger for a slow test than for a fast test. This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing).

% Frequent proactive testing of asymptomatic individuals offers an additional way to reduce transmission, and is among the most powerful and least disruptive of non-pharmaceutical interventions such as face masks, social distancing.
% \bmb{As JD suggests, the focus of the paper is not on unraveling data (although that is something we \emph{could} do, and are trying to do, with this modeling framework as well). I would say that the ``bullet points'' to start the paper are (1) the processes driving the observed dynamics of the COVID-19 epidemic include both epidemiological processes (infection, recovery, etc.) and testing processes (``processes'' is repeated here, could rephrase); (2) in addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine; (3) your last sentence in the \P\ above.}

>>>>

Here, we develop a compartmental model to incorporate epidemic processes and testing processes mechanistically to study the potential effect of testing strategies, testing intensity, test sensitivity and specificity, reporting time and isolation on epidemic dynamic. We investigate the effect of underlying parameters of the model on the basic reproduction number, $\Rnum$. While targeted testing strategies are always more effective than random testing, as expected, we find that in some cases the direct effect of testing is viral spread is expected to be stronger for a slow test than for a fast test. This counter-intuitive effect can occur when people are cautious when awaiting a test result, and may not be robust to second-order effects of fast testing (such as better contact tracing). 
\ali{discussion maybe: all of the accessible testing data is being used for one of the other three purposes rather than surveillance. These are targeted testing strategies with a different bias.}

\fady{Not sure if interesting: \\
one point we can make when discussing the relative weight of testing in different compartments is to discuss pre-testing screening tools, such as surveys or questionnaires. If we have a quantitative description of how the testing intensities affect the dynamics, we can make statements like ``our results suggest that employing a pre-testing screening tool can help target infected individuals more effectively. In particular, doubling the sensitivity of the pre-screening tool would *do something* to $\Rnum$.}
% Frequent proactive testing of asymptomatic individuals offers an additional way to reduce transmission, and is among the most powerful and least disruptive of non-pharmaceutical interventions such as face masks, social distancing.
% \bmb{As JD suggests, the focus of the paper is not on unraveling data (although that is something we \emph{could} do, and are trying to do, with this modeling framework as well). I would say that the ``bullet points'' to start the paper are (1) the processes driving the observed dynamics of the COVID-19 epidemic include both epidemiological processes (infection, recovery, etc.) and testing processes (``processes'' is repeated here, could rephrase); (2) in addition to determining what is actually observed (case reports), testing processes can also feed back to affect the epidemiological dynamics through isolation and quarantine; (3) your last sentence in the \P\ above.}

### 2021-01-17 codes/sir_macpan.Rnw
Faddy updated

startdate <- as.Date("2020-12-11")
enddate <- as.Date("2021-01-12")
## BMB: if we use the original, short time series
##   Error in stats::filter(i, kern, sides = 1) : 
##     'filter' is longer than time series

@@

## plot the resulting fit
plot(fitted.mod,data=df_covid_on)

@@

>>>>

startdate <- as.Date("2020-12-11")
enddate <- as.Date("2021-01-14")
## BMB: if we use the original, short time series
##   Error in stats::filter(i, kern, sides = 1) : 
##     'filter' is longer than time series

@@

## plot the resulting fit
plot(fitted.mod,data=df_covid_on) # all
plot(fitted.mod,data=df_covid_on, # just report data
     predict_args=list(keep_vars=c("report")))

coef(fitted.mod) ## output all fitted params
coef(fitted.mod,"fitted") ##output just beta0
## Note that that the value of \code{beta0} used for the simulation was 1

@@

# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
## Forecasting
# %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
## Here, an example involving changing the transmission rate ($\beta$) in the future due to lockdown ...

## Goal: to forcast after the date of lockdown
lockdown <- as.Date("2021-01-14")
enddate_f <- as.Date("2021-01-31") ## for forecasting
